Appendix: A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-05-02 appendix.Rmd
add allocation plot

Because $rank(\boldsymbol{R}) = n - p - 1 < n$, $\boldsymbol{e}$ follows a degenerate multivariate normal distribution and does not have a density. Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0. Then, the residuals will assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified. If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.

@@

XXX PLACEHOLDER FOR CODE FOR FIGURES

```{r}
knitr::include_graphics("figures/check_allocation_for_subject_nonlin.png")
knitr::include_graphics("figures/check_allocation_for_subject_hetero.png")
```

>>>>

Because $rank(\boldsymbol{R}) = n - p - 1 < n$, $\boldsymbol{e}$ follows a degenerate multivariate normal distribution and does not have a density. Since the Kullback-Leibler divergence requires a proper density function, we need to simplify the covariance matrix of $\boldsymbol{e}$ by setting all the off-diagonal elements to 0. Then, the residuals will be assumed to follow $N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))$ under the null hypothesis that the model is correctly specified. If the model is however misspecified due to omitted variables $\boldsymbol{Z}$, or a non-constant variance $\boldsymbol{V}$, the distribution of residuals can be derived as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))$ and $N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$ respectively.

@@

In Figure \ref{fig:poly-allocate} and Figure \ref{fig:heter-allocate}, we present examples of lineup allocations generated by our algorithm for data collection periods I and II, respectively. These figures illustrate the factor values presented to the first 20 participants recruited during each period. It can be observed that the algorithm effectively distributed almost all possible one-way and two-way factor combinations among the participants, thereby ensuring a diverse set of lineups for each participant.

```{r poly-allocate, fig.cap = 'Factor values assigned to the first 20 participants recruited during data collection period I, with data points slightly jittered to prevent overlap. Each participant have been exposed to all one-way and two-way factor combinations.'}
vi_survey %>%
  filter(exp == 3) %>%
  filter(!attention_check) %>%
  filter(set <= 20) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
  mutate(e_sigma = factor(e_sigma)) %>%
  mutate(n = factor(n)) %>%
  ggplot() +
  geom_jitter(aes(e_sigma, shape, col = n),
             width = 0.2,
             height = 0.2,
             alpha = 0.5,
             size = 3) +
  facet_wrap(~set) +
  scale_color_brewer(palette = "Dark2") +
  theme_light() +
  ylab("Non-linearity pattern") +
  xlab(expression(sigma)) +
  labs(col = "Sample size")
```

```{r heter-allocate, fig.cap = 'Factor values assigned to the first 20 participants recruited during data collection period II, with data points slightly jittered to prevent overlap. Each participant appears to have been exposed to almost all one-way and two-way factor combinations. However, two participants, namely participant 16 and participant 18, were not presented with lineups containing a butterfly shape with $b = 4$. Additionally, participant 7 did not encounter a lineup featuring a butterfly shape with $b = 1$.'}
vi_survey %>%
  filter(exp == 4) %>%
  filter(!attention_check, !null_lineup) %>%
  filter(set <= 20) %>%
  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
  mutate(b = factor(b)) %>%
  mutate(n = factor(n)) %>%
  ggplot() +
  geom_jitter(aes(b, a, col = n),
             width = 0.2,
             height = 0.2,
             alpha = 0.5,
             size = 3) +
  facet_wrap(~set) +
  scale_color_brewer(palette = "Dark2") +
  theme_light() +
  ylab("Heteroskedasticity pattern") +
  xlab("b") +
  labs(col = "Sample size")
```

### 2023-05-02 appendix.tex
use google ref for all citations
add allocation plot

Because \(rank(\boldsymbol{R}) = n - p - 1 < n\), \(\boldsymbol{e}\)
follows a degenerate multivariate normal distribution and does not have
a density. Since the Kullback-Leibler divergence requires a proper
density function, we need to simplify the covariance matrix of
\(\boldsymbol{e}\) by setting all the off-diagonal elements to 0. Then,
the residuals will assumed to follow
\(N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))\) under the null
hypothesis that the model is correctly specified. If the model is
however misspecified due to omitted variables \(\boldsymbol{Z}\), or a
non-constant variance \(\boldsymbol{V}\), the distribution of residuals
can be derived as
\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))\)
and
\(N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\)
respectively.

@@

The parameter \(\alpha\) used for the \(p\)-value calculation needs to
be estimated from responses to null lineups. With a greater value of
\(\hat{\alpha}\), the \(p\)-value will be smaller, resulting in more
lineups being rejected. However, The way we generate Rorschach lineup is
not strictly the same as what suggested in
\citet{vanderplas2021statistical} and \citet{buja_statistical_2009}.
Therefore, we conduct a sensitivity analysis in this section to examine
the impact of the variation of the estimator \(\alpha\) on our primary
findings.

@@

>>>>

Because \(rank(\boldsymbol{R}) = n - p - 1 < n\), \(\boldsymbol{e}\)
follows a degenerate multivariate normal distribution and does not have
a density. Since the Kullback-Leibler divergence requires a proper
density function, we need to simplify the covariance matrix of
\(\boldsymbol{e}\) by setting all the off-diagonal elements to 0. Then,
the residuals will be assumed to follow
\(N(\boldsymbol{0}, diag(\boldsymbol{R}\sigma^2))\) under the null
hypothesis that the model is correctly specified. If the model is
however misspecified due to omitted variables \(\boldsymbol{Z}\), or a
non-constant variance \(\boldsymbol{V}\), the distribution of residuals
can be derived as
\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\sigma^2))\)
and
\(N(\boldsymbol{0}, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\)
respectively.

@@

The parameter \(\alpha\) used for the \(p\)-value calculation needs to
be estimated from responses to null lineups. With a greater value of
\(\hat{\alpha}\), the \(p\)-value will be smaller, resulting in more
lineups being rejected. However, The way we generate Rorschach lineup is
not strictly the same as what suggested in
\citet{vanderplas2021statistical} and \citet{buja2009statistical}.
Therefore, we conduct a sensitivity analysis in this section to examine
the impact of the variation of the estimator \(\alpha\) on our primary
findings.

@@

In Figure \ref{fig:poly-allocate} and Figure \ref{fig:heter-allocate},
we present examples of lineup allocations generated by our algorithm for
data collection periods I and II, respectively. These figures illustrate
the factor values presented to the first 20 participants recruited
during each period. It can be observed that the algorithm effectively
distributed almost all possible one-way and two-way factor combinations
among the participants, thereby ensuring a diverse set of lineups for
each participant.

{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/poly-allocate-1}

\caption{Factor values assigned to the first 20 participants recruited during data collection period I, with data points slightly jittered to prevent overlap. Each participant have been exposed to all one-way and two-way factor combinations.}\label{fig:poly-allocate}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{appendix_files/figure-latex/heter-allocate-1}

\caption{Factor values assigned to the first 20 participants recruited during data collection period II, with data points slightly jittered to prevent overlap. Each participant appears to have been exposed to almost all one-way and two-way factor combinations. However, two participants, namely participant 16 and participant 18, were not presented with lineups containing a butterfly shape with $b = 4$. Additionally, participant 7 did not encounter a lineup featuring a butterfly shape with $b = 1$.}\label{fig:heter-allocate}
\end{figure}
